---
title: "Rintamaki Replication"
author: "Ebba Mark"
date: "`r Sys.Date()`"
output:
  html_document: default
---


```{css, echo=FALSE}

pre {
  max-height: 300px;
  overflow-y: auto;
}

```


```{r setup, include=FALSE}
library(tidyverse)
library(here)
library(gets)
library(readxl)
library(janitor)
library(conflicted)
library(patchwork)
library(modelsummary)
conflict_prefer_all("gets", quiet = TRUE)
conflicts_prefer(dplyr::filter)
knitr::opts_chunk$set(echo = TRUE)

source(here("code/rintamaki_replication/models/helper_code.R"))

```

# Replication {.tabset}

Recall [Rintamäki et al. 2017](https://doi.org/10.1016/j.eneco.2016.12.019) in which the authors investigate whether renewable energy generation impacts the volatility of electricity prices in Germany and Denmark (two zones in DK). They use the standard deviation of hourly prices to measure daily price volatility (and SD of daily prices for weekly volatility). They then estimate a SARMA model in each case accouning for short-term price volatility as well as weekly seasonality in the data.

*Abstract:*
*"Although variable renewable energy (VRE) technologies with zero marginal costs decrease electricity prices, the literature is inconclusive about how the resulting shift in the supply curves impacts price volatility. Because the flexibility to respond to high peak and low off-peak prices is crucial for demand-response applications and may compensate for the losses of conventional generators caused by lower average prices, there is a need to understand how the penetration of VRE affects volatility. In this paper, we build distributed lag models with Danish and German data to estimate the impact of VRE generation on electricity price volatility. We find that in Denmark wind power decreases the daily volatility of prices by flattening the hourly price profile, but in Germany it increases the volatility because it has a stronger impact on off-peak prices. Our analysis suggests that access to flexible generation capacity and wind power generation patterns contribute to these differing impacts. Meanwhile, solar power decreases price volatility in Germany. By contrast, the weekly volatility of prices increases in both areas due to the intermittency of VRE. Thus, policy measures for facilitating the integration of VRE should be tailored to such region-specific patterns."*

In each "Replicate" tab I first replicate the model from Rintamäki et al directly (excluding the control for natural gas prices as this data is proprietary/unavailable) and then make a first attempt at measuring the volatility of hourly prices directly using gets. In the "Extending to Europe" tab I provide plots of the time series of electricity price data available from EMBER (more detail on data in that section). I have chosen to include the code blocks for your reference.


## Replicate Daily 
In the specification looking at daily price volatility the dependent variable of interest is the standard deviation of hourly prices per day. We have the advantage of being able to model the volatility of hourly prices directly with gets.

### Models as in Rintamäki et al.

The below specifications replicate the results reported in Rintamaki et al for daily price volatility. Unfortunately, gas spot prices are not provided in their replication data as they are proprietary (available via Bloomberg portal) - those specs have been commented out below.

Pulled from replication data available at: <https://doi-org.ezproxy-prd.bodleian.ox.ac.uk/10.1016/j.eneco.2016.12.019>

```{r, cache = TRUE}

# Required including "here("code/rintamaki_replication/data/" to data read commands in the listed files 
source(here("code/rintamaki_replication/models/read_daily_data.R"))

```


```{r, cache = TRUE}

daily_forms <- list(
  # SPECS FOR DK1
  list(vol_dk1, wind_dk1, 2, 1, 2, 1, 7),
  list(vol_dk1, wind_pen_dk1, 2, 1, 2, 1, 7),
  list(vol_dk1, cbind(wind_dk1, exim_op1_dk1, exim_op2_dk1, exim_p_dk1), 2, 1, 2, 1, 7),
  list(vol_dk1, cbind(wind_dk1, exim_op1_dk1, exim_p_dk1), 2, 1, 2, 1, 7),
  list(vol_dk1, cbind(wind_pen_dk1, exim_op1_dk1, exim_p_dk1), 2, 1, 2, 1, 7),
  #list(fwd1(vol_dk1), list(fwd1(wind_dk1), diff(gas_dk1))),
  #SPECS FOR DK2
  list(vol_dk2, wind_dk2, 2, 1, 2, 1, 7),
  list(vol_dk2, wind_pen_dk2, 2, 1, 2, 1, 7),
  list(vol_dk2, cbind(wind_dk2, exim_op1_dk2, exim_op2_dk2, exim_p_dk2), 2, 1, 2, 1, 7),
  list(vol_dk2, cbind(wind_dk2, exim_op1_dk2, exim_p_dk2), 2, 1, 2, 1, 7),
  list(vol_dk2, cbind(wind_pen_dk2, exim_op1_dk2, exim_p_dk2), 2, 1, 2, 1, 7),
  #list(fwd1(vol_dk2), list(fwd1(wind_dk2), diff(gas_dk2))),
  # SPECS FOR DE
  list(vol_de, wind_de, 2, 1, 2, 1, 7),
  list(fwd1(vol_de), diff(solar_de), 2, 1, 2, 1, 7),
  list(fwd1(vol_de), cbind(fwd1(wind_de), diff(solar_de)), 2, 1, 2, 1, 7),
  list(vol_de, res_de, 2, 1, 2, 1, 7),
  list(vol_de, wind_pen_de, 2, 1, 2, 1, 7),
  list(fwd1(vol_de), diff(solar_pen_de), 2, 1, 2, 1, 7),
  list(vol_de, res_pen_de, 2, 1, 2, 1, 7),
  list(fwd1(vol_de), cbind(fwd1(wind_de), diff(solar_de), fwd1(exim_op1_de), fwd1(exim_op2_de), fwd1(exim_p_de)), 2, 1, 2, 1, 7))
  #list(fwd1(vol_de), list(fwd1(wind_de), diff(solar_de), diff(gas_de))))

daily_res <- lapply(daily_forms, function(x) do.call(fitModel, x))

daily_res %>% 
  modelsummary


```

### Hourly price volatility model with gets {.tabset}
This is where we have our current comparative advantage in that we can model hourly price volatility directly.

#### Data

```{r, cache = TRUE}

# Loads hourly price data for de, dk1, dk2
for(cty in c('dk1', 'dk2', 'de')){
  temp <- read_xlsx(here('code/rintamaki_replication/data/raw_data.xlsx'), sheet = cty) %>% 
    clean_names
  assign(paste0("data_", cty), temp)
}

```

#### Germany

```{r, cache = TRUE}

data_de %>% 
  ggplot(aes(x = date, y = de_price)) +
  geom_line(linewidth = 0.1) +
  labs(title = "Germany: Hourly Electricity Prices", y = "Hourly Electricity Price", x = "Date")

# Check to make sure data is ordered
data_de %>% arrange(date, hour) %>% identical(data_de)

# The below model incorporates ARCH(24) to be selected over, available regressors for wind forecasts, solar forecasts, and imports to and from France. In line with Rintamaki et al, we incorporate moving average terms for half-day, daily, and weekly prices volatility (levels?). Seasonal moving average terms are not yet incorporated here as they are in Rintamaki et al. Rather we might want to consider the seasonal moving average to have a periodicity of 24 hours (moving average over the same time of day)
de_mod <- arx(data_de$de_price, 
              # Include 24 ARCH terms to be selected over
              # Question: In theory, the ARCH(24) would proxy the seasonal (daily) AR term?
              arch = 1:24, 
              # Additional regressors to select over (only wind forecasts, solar forecasts, and imports between Germany and France are available at hourly frequency in the replication data)
              # Worth doing additional data scoping
              # Also perhaps worth including additional indicators for time of day (assuming this variation is not mopped up by the moving average terms)
              vxreg = as.matrix(data_de[c('de_wind_fcast', 'de_solar_fcast', 'de_fr', 'fr_de')]), 
              # As in Rintamaki et al. the below line incorporates moving average terms for the half-day (12), day (24 hours) and week (168 hours) 
              log.ewma = c(12, 24, 168)) 


# Perform gets model selection over the log-variance model above
# Turn of ARCH diagnostics as the GUM does not pass (could possible manipulate lag and p value choice here)
de_vmod <- getsv(de_mod, t.pval = 0.001, ar.LjungB = NULL)

de_vmod %>% plot

```

#### DK1

```{r, cache = TRUE}

data_dk1 %>% 
  ggplot(aes(x = date, y = dk1_price)) +
  geom_line(linewidth = 0.1) +
  labs(main = "DK1: Hourly Electricity Prices")

data_dk1 %>% arrange(date, hour) %>% identical(data_dk1)

# Same spec as above with appropriate corrections to the regressors
dk1_mod <- arx(data_dk1$dk1_price, 
              arch = 1:24, 
              vxreg = as.matrix(data_dk1[c("dk1_load_fcast", "dk1_wind_fcast", "dk1_no2", "no2_dk1", "dk1_se3", "se3_dk1")]),
               log.ewma = c(12, 24, 168)) 


# Perform gets model selection over the log-variance model above
dk1_vmod <- getsv(dk1_mod, t.pval = 0.001, ar.LjungB = NULL)

dk1_vmod %>% plot

```

#### DK2

```{r, cache = TRUE}

data_dk2 %>% 
  ggplot(aes(x = date, y = dk2_price)) +
  geom_line(linewidth = 0.1) +
  labs(main = "DK2: Hourly Electricity Prices")

data_dk2 %>% arrange(date, hour) %>% identical(data_dk2)

# Same spec as above with appropriate corrections to the regressors
dk2_mod <- arx(data_dk2$dk2_price, 
              arch = 1:24, 
              vxreg = as.matrix(data_dk2[c("dk2_load_fcast", "dk2_wind_fcast", 
                                 "dk2_se4","dk2_se4_2")]),
               log.ewma = c(12, 24, 168)) 


# Perform gets model selection over the log-variance model above
dk2_vmod <- getsv(dk2_mod, t.pval = 0.001, ar.LjungB = NULL)

dk2_vmod %>% plot

```


## Replicate Weekly 

Again, in the specification looking at weekly price volatility the dependent variable of interest is the standard deviation of daily prices per week. We have the advantage of being able to model the volatility of daily prices directly with gets.

### Data

Pulled from replication data available at:<https://doi-org.ezproxy-prd.bodleian.ox.ac.uk/10.1016/j.eneco.2016.12.019>

```{r, cache = TRUE}

# Required including "here("code/rintamaki_replication/data/" to data read commands in the listed files
source(here("code/rintamaki_replication/models/read_weekly_data.R"))

```

### Models as in Rintamäki et al.
```{r, cache = TRUE}

weekly_forms <- list(
  # SPECS FOR DK1
  list(vol_dk1, wind_dk1, 1, 0, 0, 1, 4),
  list(fwd1(vol_dk1), cbind(fwd1(wind_dk1), diff(exim_dk1)), 1, 0, 0, 1, 4),
  list(vol_dk1, wind_std_dk1, 1, 0, 0, 1, 4),
  # SPECS FOR DK2
  list(vol_dk2, wind_dk2, 1, 0, 0, 1, 4),
  list(vol_dk2, wind_std_dk2, 1, 0, 0, 1, 4),
  # SPECS FOR DE
  list(fwd1(vol_de), diff(wind_de), 1, 0, 0, 0, 4),
  list(fwd1(vol_de), diff(solar_de), 1, 0, 0, 0, 4),
  list(vol_de, res_de, 1, 0, 0, 0, 4),
  list(fwd1(vol_de), cbind(diff(wind_de), fwd1(exim_de)), 1, 0, 0, 0, 4))

weekly_res <- lapply(weekly_forms, function(x) do.call(fitModel, x))

weekly_res %>%
  modelsummary


```

### Daily price volatility model with gets {.tabset}

#### Data

```{r, cache = TRUE}

for(cty in c('dk1', 'dk2', 'de')){
  temp <- read_xlsx(here('code/rintamaki_replication/data/raw_data.xlsx'), sheet = cty) %>%
    clean_names
  assign(paste0("data_", cty), temp)
}

df_list <- list("data_de" = data_de, "data_dk1" = data_dk1, "data_dk2" = data_dk2)

for(n_df in names(df_list)){
  df <- df_list[[n_df]]
  temp <- df %>%
    tibble %>%
    # Certain days in October have multiple values for hour one across all countries...not sure why.
    # Below I take the average of the duplicate observations but might be worth looking into why this is the case
    group_by(date, hour) %>%
    summarise(across(names(df)[!(names(df) %in% c("date", "hour", "month"))], ~mean(., na.rm = TRUE))) %>%
    # Daily average price
    group_by(date) %>%
    select(-c("hour")) %>%
    summarise(across(names(df)[!(names(df) %in% c("date", "hour", "month"))], ~mean(., na.rm = TRUE)))

   assign(paste0("daily_", n_df), temp)
}

```

#### Germany

```{r, cache = TRUE}

daily_data_de %>%
  ggplot(aes(x = date, y = de_price)) +
  geom_line(linewidth = 0.1) +
  labs(main = "Germany: Daily Electricity Prices")

daily_data_de %>% arrange(date) %>% identical(daily_data_de)

# equally-weighted moving averages are incorporated to account for multi-day moving averages 2-, 5-, 7-, 14-day moving averages

# Might want to consider re-incorporating the seasonality dimension as well
daily_de_mod <- arx(daily_data_de$de_price,
              # Include 14 ARCH terms to be selected over (two-week period)
              # in theory, the ARHC(14) would proxy the seasonal AR term?
              arch = 1:14,
              # Additional regressors to select over (only wind forecasts, solar forecasts, and imports between Germany and France are available at hourly frequency in the replication data
              # Worth doing additional data scoping
              # Also perhaps worth including additional indicators for time of day (assuming this variation is not mopped up by the moving average terms)
              vxreg = as.matrix(daily_data_de[c('de_load_fcast', 'de_wind_fcast', 'de_solar_fcast', 'de_fr', 'fr_de')]),
              # As in Rintamaki et al. the below line incorporates moving average terms for the two-day, 5-day, one-week, and two-week prices
              log.ewma = c(2, 5, 7, 14))
# Seasonal moving average terms are not yet incorporated here as they are in Rintamaki et al. Rather we might want to consider the seasonal moving average to have a periodicity of X hours (moving average over the same time of day)

# Perform gets model selection over the log-variance model above
daily_de_vmod <- getsv(daily_de_mod, t.pval = 0.001, ar.LjungB = NULL)

daily_de_vmod %>% plot

```

#### DK1

```{r, cache = TRUE}

daily_data_dk1 %>%
  ggplot(aes(x = date, y = dk1_price)) +
  geom_line(linewidth = 0.1) +
  labs(main = "DK1: Daily Electricity Prices")

daily_data_dk1 %>% arrange(date) %>% identical(daily_data_dk1)

# Same spec as above with appropriate corrections to the regressors
daily_dk1_mod <- arx(daily_data_dk1$dk1_price,
              mc = TRUE,
              arch = 1:14,
              vxreg = as.matrix(daily_data_dk1[c("dk1_load_fcast", "dk1_wind_fcast", "dk1_no2", "no2_dk1", "dk1_se3", "se3_dk1")]),
               log.ewma = c(2, 5, 7, 14))


# Perform gets model selection over the log-variance model above
daily_dk1_vmod <- getsv(daily_dk1_mod, t.pval = 0.001, ar.LjungB = NULL)

daily_dk1_vmod %>% plot

```

#### DK2

```{r, cache = TRUE}

daily_data_dk2 %>%
  ggplot(aes(x = date, y = dk2_price)) +
  geom_line(linewidth = 0.1) +
  labs(main = "DK2: Daily Electricity Prices")

data_dk2 %>% arrange(date) %>% identical(data_dk2)

# Same spec as above with appropriate corrections to the regressors
daily_dk2_mod <- arx(daily_data_dk2$dk2_price,
              arch = 1:14,
              vxreg = as.matrix(daily_data_dk2[c("dk2_load_fcast", "dk2_wind_fcast",
                                 "dk2_se4","dk2_se4_2")]),
               log.ewma = c(2, 5, 7, 14))


# Perform gets model selection over the log-variance model above
# Requires turning off both ARCH and AR Ljung-Box test - could play around with lag and pvalues
daily_dk2_vmod <- getsv(daily_dk2_mod, t.pval = 0.001, ar.LjungB = NULL, arch.LjungB = NULL)

daily_dk2_vmod %>% plot

```


## Replicate Intraday

Pulled from replication data available at:<https://doi-org.ezproxy-prd.bodleian.ox.ac.uk/10.1016/j.eneco.2016.12.019>

```{r, cache = TRUE}

# Required including "here("code/rintamaki_replication/data/" to data read commands in the listed files 
source(here("code/rintamaki_replication/models/read_intraday_data.R"))

```

### Models as in Rintamäki et al.
```{r, cache = TRUE}

intraday_forms <- list(
  list(price_op1_dk1, cbind(wind_op1_dk1, exim_op1_dk1), 2, 1, 1, 1, 7),
  list(price_p_dk1, cbind(wind_p_dk1, exim_p_dk1), 2, 1, 1, 1, 7),
  list(price_op2_dk1, cbind(wind_op2_dk1, exim_op2_dk1), 2, 1, 1, 1, 7),
  list(price_p_dk2, cbind(wind_p_dk2, exim_p_dk2),  2, 1, 1, 1, 7),
  list(price_op2_dk2, cbind(wind_op2_dk2, exim_op2_dk2),  2, 1, 1, 1, 7),
  list(price_op1_de, cbind(wind_op1_de, exim_op1_de),  2, 1, 1, 1, 7),
  list(price_p_de, cbind(wind_p_de, exim_p_de), 2, 1, 1, 1, 7),
  list(price_op2_de, cbind(wind_op2_de, exim_op2_de),  2, 1, 1, 1, 7),
  list(fwd1(price_p_de), cbind(diff(solar_p_de), fwd1(exim_p_de)),  2, 1, 1, 1, 7), 
  list(price_p_de, cbind(res_p_de, exim_p_de), 2, 1, 1, 1, 7))


intraday_res <- lapply(intraday_forms, function(x) do.call(fitModel, x))

modelsummary(intraday_res)

```

### Intraday price model with gets  {.tabset}

#### Data

```{r, cache = TRUE}

for(cty in c("de","dk1", "dk2")){
  temp <- read.table(here(paste0("code/rintamaki_replication/data/", cty, "_intraday.txt")), header=TRUE) %>% 
  tibble %>% 
  mutate(date = dmy(date)) %>% 
  arrange(date)
  
  assign(paste0("intraday_data_", cty), temp)
}
```

#### Germany

```{r, cache = TRUE}

intraday_data_de %>% 
  select(date, op1_price, op2_price, p_price) %>% 
  pivot_longer(cols = !date) %>% 
  ggplot() +
  geom_line(aes(x = date, y = value, color = name), linewidth = 0.1) +
  labs(main = "DE: Intraday Electricity Prices")

intraday_data_de %>% arrange(date) %>% identical(intraday_data_de)

for(dep in c("op1_", "op2_", "p_")){
  temp <- intraday_data_de %>% 
    select(date, contains(dep))
  print(temp)
  
  dep2 <- as.matrix(temp[paste0(dep, "price")])
  vxregs <- as.matrix(temp[!(names(temp) %in% c("date", paste0(dep, "price")))])

  temp_mod <- arx(dep2, 
              ar = 1:14, 
              vxreg = vxregs,
              log.ewma = c(2, 5, 7, 14)) 
  assign(paste0("de_", dep, "mod"), temp_mod)
  
  temp_vmod <- getsv(temp_mod, t.pval = 0.001, ar.LjungB = NULL, arch.LjungB = NULL)
  assign(paste0("de_", dep, "vmod"), temp_vmod)

  plot(temp_vmod)
  
}

```

#### DK1

```{r, cache = TRUE}

intraday_data_dk1 %>% 
  select(date, op1_dk1_price, op2_dk1_price, p_dk1_price) %>% 
  pivot_longer(cols = !date) %>% 
  ggplot() +
  geom_line(aes(x = date, y = value, color = name), linewidth = 0.1) +
  labs(main = "DK1: Intraday Electricity Prices")

intraday_data_dk1 %>% arrange(date) %>% identical(intraday_data_dk1)

for(dep in c("op1_dk1_", "op2_dk1_", "p_dk1_")){
  temp <- intraday_data_dk1 %>% 
    select(date, contains(dep))
  print(temp)
  
  dep2 <- as.matrix(temp[paste0(dep, "price")])
  vxregs <- as.matrix(temp[!(names(temp) %in% c("date", paste0(dep, "price")))])

  temp_mod <- arx(dep2, 
              arch = 1:14, 
              vxreg = vxregs,
              log.ewma = c(2, 5, 7, 14)) 
  assign(paste0("dk1_", dep, "mod"), temp_mod)
  
  temp_vmod <- getsv(temp_mod, t.pval = 0.001, ar.LjungB = NULL, arch.LjungB = NULL)
  assign(paste0("dk1_", dep, "vmod"), temp_vmod)

  plot(temp_vmod)
  
}

```

#### DK2

```{r, cache = TRUE}

intraday_data_dk2 %>% 
  select(date, op1_dk2_price, op2_dk2_price, p_dk2_price) %>% 
  pivot_longer(cols = !date) %>% 
  ggplot() +
  geom_line(aes(x = date, y = value, color = name), linewidth = 0.1) +
  labs(main = "dk2: Intraday Electricity Prices")

intraday_data_dk2 %>% arrange(date) %>% identical(intraday_data_dk2)

for(dep in c("op1_dk2_", "op2_dk2_", "p_dk2_")){
  temp <- intraday_data_dk2 %>% 
    select(date, contains(dep))
  print(temp)
  
  dep2 <- as.matrix(temp[paste0(dep, "price")])
  vxregs <- as.matrix(temp[!(names(temp) %in% c("date", paste0(dep, "price")))])

  temp_mod <- arx(dep2, 
              arch = 1:14, 
              vxreg = vxregs,
              log.ewma = c(2, 5, 7, 14)) 
  assign(paste0("dk2_", dep, "mod"), temp_mod)
  
  temp_vmod <- getsv(temp_mod, t.pval = 0.001, ar.LjungB = NULL, arch.LjungB = NULL)
  assign(paste0("dk2_", dep, "vmod"), temp_vmod)

  printtex(temp_vmod)
  plot(temp_vmod)
  
}


```

## Extending to Europe {.tabset}

### Hourly prices

Using data from [EMBER](https://ember-energy.org/data/european-wholesale-electricity-price-data/) on European wholesale electricity prices (hourly - daily and monthly also exist), I perform the same analysis as above. The hourly prices are the day-ahead wholesale prices by country.

```{r, echo = FALSE, fig.height = 14}

eu_hourly <- read.csv(here('data/raw/EMBER_europe_wholesale_data/all_countries.csv')) %>% 
  tibble %>% 
  clean_names %>% 
  mutate(across(contains("datetime"), ~ymd_hms(.x)))

eu_hourly %>% 
  ggplot(aes(x = datetime_local, y = price_eur_m_whe)) +
  geom_line() +
  facet_wrap(~country, ncol = 4) +
  labs(title = "Hourly Wholesale Electricity Prices", subtitle = "Prices are average day-ahead spot prices per MWh sold per time period", x = "Date", y = "Euro/MWh") +
  theme_minimal()
  
```


#### Spotlight on UK, Germany, Denmark

Here, see the time series for Germany and Denmark (to compare to the data avialable in Rintamaki et al. - color portion of the TS plots). UK spotlight to discuss potential interest to David. 

```{r, echo = FALSE, fig.height = 10}

# Germany
p_uk <- eu_hourly %>% 
  filter(country == "United Kingdom") %>% 
  ggplot() +
  geom_line(aes(x = datetime_local, y = price_eur_m_whe), linewidth = 0.1) +
  labs(title = "UK: Hourly Electricity Prices", y = "Hourly Electricity Price (EUR/MWH)", x = "Date")

# Germany
p_de <- data_de %>% 
  ggplot(aes(x = date, y = de_price)) +
  geom_line(linewidth = 0.1, color = "purple") +
  geom_line(data = filter(eu_hourly, country == "Germany"), aes(x = datetime_local, y = price_eur_m_whe), linewidth = 0.1) +
  labs(title = "Germany: Hourly Electricity Prices", y = "Hourly Electricity Price (EUR/MWH)", x = "Date")

# DK1
p_dk1 <- ggplot() +
  geom_line(data = data_dk1, aes(x = date, y = dk1_price), linewidth = 0.1, color = "darkgreen") +
  geom_line(data = filter(eu_hourly, country == "Denmark"), aes(x = datetime_local, y = price_eur_m_whe)) +
  labs(title = "Denmark: Hourly Electricity Prices", subtitle = "Using DK1 prices from Rintamaki et al.", y = "Hourly Electricity Price (EUR/MWH)", x = "Date")

# DK2
p_dk2 <- data_dk2 %>% 
  ggplot(aes(x = date, y = dk2_price)) +
  geom_line(linewidth = 0.1, color = "coral") +
  geom_line(data = filter(eu_hourly, country == "Denmark"), aes(x = datetime_local, y = price_eur_m_whe)) +
  labs(title = "Denmark: Hourly Electricity Prices", subtitle = "Using DK2 prices from Rintamaki et al.", y = "Hourly Electricity Price (EUR/MWH)", x = "Date")


((p_uk + p_de) / (p_dk1 + p_dk2)) + plot_annotation(
  title = 'Hourly Wholesale Electricity Price Data UK, Germany, and Denmark',
  subtitle = 'Black represents EMBER data, colorful data represents data from Rintamaki replication.',

)

```


### Daily and Monthly Prices
```{r, echo = FALSE, fig.height = 12}

eu_daily <- read.csv(here('data/raw/EMBER_europe_wholesale_data/european_wholesale_electricity_price_data_daily.csv')) %>% 
  tibble %>% 
  clean_names %>% 
  mutate(date = ymd(date))


eu_monthly <- read.csv(here('data/raw/EMBER_europe_wholesale_data/european_wholesale_electricity_price_data_monthly.csv')) %>% 
  tibble %>% 
  clean_names %>% 
  mutate(date = ymd(date))

ggplot() +
  geom_line(data = eu_daily, aes(x = date, y = price_eur_m_whe)) +
  geom_line(data = eu_monthly, aes(x = date, y = price_eur_m_whe), color = "steelblue") +
  facet_wrap(~country) +
  labs(title = "Daily and Monthly Wholesale Electricity Prices", subtitle = "Black (blue) represents daily (monthly) prices.", y = "Price (EUR/MWH)", x = "Date") +
  theme_minimal()

```


